Variants on the Flight and Maintenance Planning problem

### 2019-04-15 latex/notes/bounds/bounds.tex
passed bounds document into latex.

\renewcommand\subsection{\@startsection{subsection}{2}{\z@}%
                       {-6\p@ \@plus -0\p@ \@minus -0\p@}%
                       {2\p@ \@plus 0\p@ \@minus 0\p@}%
                       {\normalsize\textbf}}

\renewcommand\subsubsection{\@startsection{subsubsection}{3}{\z@}%
                       {-6\p@ \@plus -0\p@ \@minus -0\p@}%
                       {1\p@ \@plus 0\p@ \@minus 0\p@}%
                       {\normalsize\itshape\bfseries}}
\makeatother

@@

\section{Mathematical formulation}
  \label{sec:model}

The following model provides a tight MIP formulation that solves the Military Flight and Maintenance Problem. Maintenance operations will be referred as "checks".

\subsection{Input data}
    \subsubsection{Basic sets}

@@

    \subsubsection{Mission parameters}

@@

    \subsubsection{Maintenance parameters}

@@

    \subsubsection{Fleet parameters}

@@

    \subsubsection{Parametric sets}

@@

    \subsubsection{Time-related parametric sets}

@@

        \begin{tabular}{p{30mm}p{125mm}}
            $t \in \mathcal{T}^s_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}a). \\
            % $\mathcal{T}^S_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t},  ..., \min{\{|\mathcal{T}|, t + M-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
            $t' \in \mathcal{T}^m_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + M + E^{min}-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
            $t' \in \mathcal{T}^M_t$ &  time periods $t' \leq |\mathcal{T}| - E^{max} - M$ such that $t' \in \{ t + M + E^{min}-1 , ...,  t + M + E^{max}-1 \}$ (figure \ref{fig:gantt_windows}a). \\
            $t \in \mathcal{T}^{m_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}}\}$ (figure \ref{fig:gantt_windows}b). \\
            $t \in \mathcal{T}^{M_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}} , ...,  Rct^{Init}_i \}$ (figure \ref{fig:gantt_windows}b). \\
            $t' \in \mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}c). \\
            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}$ & pairs of time periods $t_1, t_2 \in \mathcal{TJ}_j$ such that $t_2 \ge t_1 + MT_j-1$ and $t \in \{t_1, ..., t_2\}$. \\
            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}$ & pairs of time periods $t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}$ such that $t_1 \in \mathcal{T}^s_{t} \lor t_2 \in \mathcal{T}^s_{t}$. \\
            $(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}$ & triplet composed of mission $j \in \mathcal{JI}_i$ and $t, t' \in \mathcal{TJ}_j$ such that $t' \ge t + MT_j-1$ and $t \ge t_1 + M$ and $t' < t_2$ \\
        \end{tabular}

    \begin{figure}
        \centering
        % \includegraphics[width=\linewidth]{img/gantt_timewindows_all.png}
        \begin{tikzpicture}
            \input{gantts/time_windows.tex}
        \end{tikzpicture}
        \begin{tikzpicture}
            \input{gantts/time_windows2.tex}
        \end{tikzpicture}
        \begin{tikzpicture}
            \input{gantts/time_windows_init.tex}
        \end{tikzpicture}
        \begin{tikzpicture}
            \input{gantts/time_windows_mission.tex}
        \end{tikzpicture}
        \caption{Examples showing the maintenance-related time-parametric sets for aircraft $i$. (a) $\mathcal{T}_{t'}^{s}$ refers to the previous $M$ periods to period $t'$. $\mathcal{T}_{t}^{m}$ refers to the periods where a check cannot be planned after starting a check in period $t$. Finally, $\mathcal{T}_{t}^{M}$ refers to the periods where a check needs to be scheduled, after starting a check in period $t$. (b) The maintenance assignment for aicraft $i$ is $m_{it_1t_2}=1$. $\mathcal{T}_A \cup \mathcal{T}_B$ represents the set of periods $t$ for which $(t_1, t_2) \in \mathcal{TTT}_t$. (c) $Rct_{i}^{Init}$ is equal to 55, meaning a check has to be planned between months number 25 and 54. The maintenance assignment for aircraft $i$ is $m_{it_1t_2}=1$ and, since it includes the last period, it only represents one operation. (d) $\mathcal{T}_{jt}^{MT}$ refers to the periods where the assignment of mission $j$ needs to be kept. In this case, the size is 6. The mission assignments for aircraft $i$ to mission $j$ are $a_{ijt_1t_2}=1$ and $a_{ijt'_1t'_2}=1$}
        \label{fig:gantt_windows}
    \end{figure}

    \subsubsection{Consumption derived parameters}\label{subsubsec:consumption}

    The following parameters are derived from the basic input parameters. They are used in constraints (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_rem_hours3}). They are transformations of the default and mission flight hours consumption under different time circumstances.

    \begin{align}
        & U^{min}_{t_1t_2} = U^{min} \sum_{t=t_1}^{t_2} (t_2 - t) \notag \\
        & H'_{jt_1t_2} = (H_j - U^{min}) (t_2 - t_1) \notag \\
        & H''_{jt_1t_2t'} = \sum_{t=t_1}^{t_2} (H_j- U^{min}) (t' - t) \notag
    \end{align}

\subsection{Variables}

    The following decision variables control the assignment of missions and checks to aircraft.

    \begin{tabular}{p{8mm}p{147mm}}
        $a_{ijtt'}$ &  =1 if aircraft $i$ starts a new assignment to mission $j$ at the beginning of period $t$ and finishes it at the end of period $t'$. \\  
        $m_{itt'}$  &  =1 if aircraft $i \in I$ starts a check at the beginning of period $t \in \mathcal{T}$ and then starts the next one at the beginning of period $t'$, 0 otherwise. \\
        $h_{itt'}$  &  Continuous. Sum of the monthly remaining flight hours for aircraft $i \in I$ for periods between the beginning of period $t$ and the beginning of period $t'$. \\
    \end{tabular}

    \paragraph{Fixed values}

    Note that $a_{ijtt'}$ and $m_{itt'}$ are initially set up to 0 for all aircraft already in maintenance at the beginning of the planning horizon for the remaining time periods of the check. $N_{t}$ is calculated based on this information.
Similarly, for aircraft that have not yet complied with their minimum mission assignment duration at the beginning of the planning horizon, $a_{ijtt'}$ is fixed to comply with the constraints.

\subsection{Objective function and constraints}

    Objective (\ref{eq:objective1}) minimizes the number of checks while at the same time maximizing the monthly total flight hours potential of the fleet.

    \begin{align}
        & \text{Min}\; \sum_{\substack{i \in \mathcal{I}, \\ t_1 \in \mathcal{T}^{M_{Init}}_i, \\ t_2 \in \mathcal{T}^M_{t_1} - |\mathcal{T}|}} m_{it_1t_2} \times H^{max} \times (t_2 - t_1) - h_{it_1t_2}  \label{eq:objective1}
    \end{align}

    The first term counts all the remaining flight hours given to aircraft following checks and the second term quantifies the amount of remaining flight hours for all aircraft at each period. These two objectives have the same units, can be easily compared and ensure the aircraft are used in the most efficient way.

    The following constraints are used in the model:
    \begin{align}
        % maximum capacity1:
        & \sum_{\substack{i \in \mathcal{I}, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + N_t \leq C^{max}
          & t \in \mathcal{T} \label{eq:capacity1}\\
        % min assignments:
        & \sum_{\substack{i \in \mathcal{IJ}_j, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}}} a_{ijt_1t_2} \geq R_j
                & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
        % just doing one thing at any given time:
        & \sum_{\substack{(t_1, t_2) \in \\ \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + \sum_{\substack{j \in \\ \mathcal{JT}_t \cap \mathcal{JI}_i}} \sum_{\substack{(t_1, t_2) \in \\ \mathcal{T}\mathcal{T}\mathcal{J}_{jt}}} a_{ijt_1t_2} \leq 1 
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
    \end{align}

    Maintenance capacity is controlled by (\ref{eq:capacity1}). The aircraft requirements of missions are defined by (\ref{eq:missionres}). Constraints (\ref{eq:state}) ensure that an aircraft can only be used for one mission or undergo check in the same period.

    \begin{align}
       & \sum_{\substack{i \in \mathcal{IK}_k, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + N^{Clust}_{kt} \leq A^{Clust}_{kt}
        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster}
       % & \sum_{i \in \mathcal{IK}_k} rft_{it} \geq H^{Clust}_{kt}
       %  &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:sustainability-cluster}
    \end{align}

    Constraints (\ref{eq:serviceability-cluster}) guarantee a minimum serviceability of aircraft for each cluster $k$. A cluster is defined by the largest group of aircraft that is required exclusively for at least one mission. 
    % Constraints (\ref{eq:sustainability-cluster}) ensure there is a minimum amount of remaining flight time for each cluster $k$.

    \begin{align}
        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H'_{jtt'} + U^{min}_{1t_1} \leq Rft^{Init}_i + H^{max} (1 - m_{it_1t_2}) 
                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours1}\\
        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H'_{jtt'} + U^{min}_{t_1t_2} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours2}\\
        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H'_{jtt'} + U^{min}_{t_2|\mathcal{T}|} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours3}
    \end{align}

        Constraints (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_hours3}) limit the total flight hours of any given aircraft between checks. Constraints (\ref{eq:cycle_hours1}) takes into account periods since the beginning of the planning horizon until the first check, \ref{eq:cycle_hours2} takes into account periods between the first and second check and \ref{eq:cycle_hours3} takes into account periods between the second check and the end of the planning horizon.

    \begin{align}
        & h_{i1t_1} \geq (t_1 - 1) Rft^{Init}_i m_{it_1t_2} - \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H''_{jtt't_1} - U^{min}_{1t_1} 
            & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours1}\\
        & h_{it_1t_2} \geq (t_2 - t_1) H^{max} m_{it_1t_2} - \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H''_{jtt't_2} - U^{min}_{t_1t_2}
            &t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours2}\\
        & h_{it_2|\mathcal{T}|} \geq (|\mathcal{T}|- t_2) H^{max} m_{it_1t_2} - \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H''_{jtt'|\mathcal{T}|} - U^{min}_{t_2|\mathcal{T}|}
            & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours3}
    \end{align}

    Constraints (\ref{eq:cycle_rem_hours1}) - (\ref{eq:cycle_rem_hours3}) define the sum of remaining flight hours each aircraft has in between checks. Similarly to (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_hours3}), their difference lies in the three possible intervals produced by assigning at most two checks. The derived parameters used in these constraints can be found in section \ref{subsubsec:consumption}.

    \begin{align}
       % For now, we only want one or possibly two maintenances...
        & \sum_{\substack{t_1 \in \mathcal{T}^{M_{Init}}_i, \\ t_2 \in \mathcal{T}^M_{t_1} \cup |\mathcal{T}|}} m_{it_1t_2} =  1 
          & i \in \mathcal{I}\label{eq:num_maint} \\
        % & m_{it'} + m_{it} \leq 1
        %   & t \in \mathcal{T}, t' \in \mathcal{T}^m_t, i \in \mathcal{I}\label{eq:rct_min}\\  
        % & \sum_{t' \in \mathcal{T}^M_t} m_{it'} \geq  m_{it}
        %   & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rct_max}\\
        % & m_{it_1t_2} = 0
        %   & t_1 \in \mathcal{T}^{m_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\}, i \in \mathcal{I}\label{eq:rct_min_init} \\
        & m_{it_1t_2} =  0
          & t_1 \notin \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}, i \in \mathcal{I} \label{eq:maint_forbid} \\
        & m_{it_1t_2} =  0
          & t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \notin \mathcal{T}^M_{t_1} \cup |\mathcal{T}|, i \in \mathcal{I} \label{eq:maint_forbid2}
    \end{align}

    Constraints (\ref{eq:num_maint}) limit the number of maintenance assignments per aircraft to one. This implies that each aircraft will get strictly one or two maintenances over the whole planning horizon.

    Constraints (\ref{eq:maint_forbid}) - (\ref{eq:maint_forbid2}) limit the combinations of maintenance schedules that can be done during the horizon assuming that the first maintenance is dependant in the initial state of the aircraft and the second maintenance (if any) is limited by the moment at which the first maintenance was planned.

\end{document}

>>>>

\renewcommand\section{\@startsection{section}{2}{\z@}%
                       {-6\p@ \@plus -0\p@ \@minus -0\p@}%
                       {2\p@ \@plus 0\p@ \@minus 0\p@}%
                       {\normalsize\textbf}}

\renewcommand\section{\@startsection{section}{3}{\z@}%
                       {-6\p@ \@plus -0\p@ \@minus -0\p@}%
                       {1\p@ \@plus 0\p@ \@minus 0\p@}%
                       {\normalsize\itshape\bfseries}}
\makeatother

@@

The idea is to define ways with which we can understand the problem, based on the initial input data.

\section{Input data}
    \subsection{Basic sets}

@@

    \subsection{Mission parameters}

@@

    \subsection{Maintenance parameters}

@@

    \subsection{Fleet parameters}

@@

    \subsection{Parametric sets}

@@

    \subsection{Time-related parametric sets}

@@

        \begin{tabular}{p{30mm}p{125mm}}
            $t \in \mathcal{T}^s_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}a). \\
            $t' \in \mathcal{T}^m_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + M + E^{min}-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
            $t' \in \mathcal{T}^M_t$ &  time periods $t' \leq |\mathcal{T}| - E^{max} - M$ such that $t' \in \{ t + M + E^{min}-1 , ...,  t + M + E^{max}-1 \}$ (figure \ref{fig:gantt_windows}a). \\
            $t \in \mathcal{T}^{m_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}}\}$ (figure \ref{fig:gantt_windows}b). \\
            $t \in \mathcal{T}^{M_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}} , ...,  Rct^{Init}_i \}$ (figure \ref{fig:gantt_windows}b). \\
            $t' \in \mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}c). \\
            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}$ & pairs of time periods $t_1, t_2 \in \mathcal{TJ}_j$ such that $t_2 \ge t_1 + MT_j-1$ and $t \in \{t_1, ..., t_2\}$. \\
            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}$ & pairs of time periods $t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}$ such that $t_1 \in \mathcal{T}^s_{t} \lor t_2 \in \mathcal{T}^s_{t}$. \\
            $(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}$ & triplet composed of mission $j \in \mathcal{JI}_i$ and $t, t' \in \mathcal{TJ}_j$ such that $t' \ge t + MT_j-1$ and $t \ge t_1 + M$ and $t' < t_2$ \\
        \end{tabular}

% \section{Number and possible dates for maintenances}
% \label{number-and-possible-dates-for-maintenances}

% The aircraft initial state already conditions the first maintenance of each aircraft.

% Moreover, using the maximum and minimum distances between maintenances (min and max size for maintenances cycles respectively), \textbf{we can precalculate the possible total number of maintenances that an aircraft will have in the horizon}. This number will probably be one of the following: {[}1{]}, {[}1, 2{]}, {[}2{]}, {[}1, 2, 3{]}. This will depend on the size of the horizon and maintenance conditions.

\section{Range of number of maintenances per period and cluster}
\label{range-of-number-of-maintenances-per-period-and-cluster}

First, for each period $t$, and using the initial states of aircraft, we can calculate the minimum and maximum number of maintenances that an aircraft could have already done (\emph{ended}) before the start of the period. We can then aggregate this for each aircraft type to obtain a lower and an upper bound on the number of maintenances that can be finished before period $t$.

\begin{align}

	& t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \forall i \in \mathcal{I}, |\mathcal{T}|

\end{align}

Secondly, we can calculate the number of maintenances we can fit until period $t$, by looking at missions requirements in terms of number of aircraft. This would be done by aircraft type. We first take the sum of all the non-mission periods over all aircraft $i \in IK_k \forall k \in K$\$ and all periods $t' < t$. Then, we divide this over the duration of the maintenances ($M$). This would an upper bound.

 Thirdly, we can sum all flight-hour needs of missions of type $k$ before time $t$. If we substract this from the initial remaining flight time of the group of aircraft, we are left with a number of hours that need to be obtained via maintenances. If we divide this number over the $H^{max}$ flight hours a maintenance gives, we get a lower bound on the number of maintenances we need to do for those aircraft of type $k$.  So, in this way, we have two lower bounds and two upper bounds per period and aircraft type. We could just get the maximum and the minimum respectively to get some bounds on the number of maintenances until period $t$.  {%

\section{Mission assignment}\label{mission-assignment}}  {%

\section{Mission assignment for initial periods}\label{mission-assignment-for-initial-periods}}

Specially for the initial periods, and based on the initial state, missions' hourly requirements and minimum assignment duration, it is possible to \textbf{precalculate which aircraft will not be able to start a mission in some periods}. We can do this only for the initial periods because we do not have the actual consumption for individual aircraft.  In the case of having a minimum consumption, this could be used to have some lower bounds too. 
\end{document}
